---
title: "EDA - Histograms"
author: "Khaled Sobeh"
date: "5/5/2022"
output: html_document
---

Data Visualization: Histograms / continuous data types 


Histogram is a kind of graphs that we use to represent a continuous variable.
1. By saying continuous, we mean a variable that can any value in an interval(such as: height, acceleration, length, time, temperature, age...), it's a quantitative variable that we can summarize.

Data structure in R: 
1. integer.
2. numeric.

Histograms
with histograms we can check if the "pattern" of a variable ,if the variable is: "symmetric", "skewed left" or "right", "unimodal"(normally distributed), "bimodal" or "multimodal". 


I will try to use only ggplot2 functions, except if the alternative is much easier, with much less code to write. 
note: plot_grid() function will be used a lot. I used it just to combine plots, as we shall see.

So, let's start.

Loading the libraries, and getting the data.
```{r, message = FALSE, warning = FALSE}
library(ggplot2)
library(priceR)
library(dplyr)
library(forcats)
library(cowplot)
library(ISLR)
```

```{r}
auto <- Auto
auto$origin <- as.factor(auto$origin)
auto$cylinders <- as.factor(auto$cylinders)
auto <- select(auto, -name)
```
```{r}
str(auto)
```

The variables that I used from the dataset:
1) acceleration: continuous variable.

## a reminder.
If you forget one day how to make plot with ggplot2(Please don't...), then you have to remember that you have to pass these three:
1) ggplot(and_here_the_dataset_name)
2) aes(x = the_variable_name_that_you_want_to_visualize, y = sec_col...)
you can pass this aes thing(you should remember this) in ggplot() or geom_..() function.
3) geom_ the_plot_that_you_want: geom_bar(), geom_boxplot().....
so you should have:
ggplot(travel,aes(acceleration)) + geom_bar() |or| ggplot(travel) + geom_bar(aes(acceleration))




## The most basic histogram:
```{r}
ggplot(auto,aes(acceleration)) + geom_histogram()
```

we can also set the aesthetic in the geom function:
```{r}
ggplot(auto) + geom_histogram(aes(acceleration))
```

```{r, message=FALSE}
plot_grid(ggplot(auto,aes(acceleration)) + geom_histogram(),
          ggplot(auto) + geom_histogram(aes(acceleration)))
```


### bins and binwidth
bins control the number of bins('bars')
binwidth control the width of the bins by x axis

more clear example: a contiuous variable: values between 1 and 100([1,100]):
```{r}
set.seed(1)
df <- data.frame(col_1 = purrr::rdunif(1000,1,100))
```

```{r}
plot_grid(ggplot(df,aes(col_1)) + geom_histogram(bins = 5, fill = "lightblue", colour = "black") +
            ggtitle('bins = 5') + scale_x_continuous(),
          ggplot(df,aes(col_1)) + geom_histogram(binwidth = 20, fill = "lightblue", colour = "black") +
            ggtitle('binwidth = 5'))
```


geom_histogram parameters:
```{r, fig.width=10,fig.height=11, message = FALSE}

plot_grid(ggplot(auto, aes(acceleration)) + geom_histogram(bins = 10) + ggtitle("bins = 10"),
          ggplot(auto, aes(acceleration)) + geom_histogram(bins = 50) + ggtitle("bins = 50"),
          ggplot(auto, aes(acceleration)) + geom_histogram(binwidth = 2) + ggtitle("binwidth = 2"),
          ggplot(auto, aes(acceleration)) + geom_histogram(binwidth = 7) + ggtitle("binwidth = 7"),
          ggplot(auto, aes(acceleration)) + geom_histogram(alpha = 0.3) + ggtitle("alpha = 0.3"),
          ggplot(auto, aes(acceleration)) + geom_histogram(alpha = 1) + ggtitle("alpha = 1"),
          ggplot(auto, aes(acceleration)) + geom_histogram(color = 'lightblue') + 
            ggtitle("color = 'lightblue'"),
          ggplot(auto, aes(acceleration)) + geom_histogram(color = '#FFDB6D') + ggtitle("color = #FFDB6D"),
          ggplot(auto, aes(acceleration)) + geom_histogram(fill = 'lightblue') + ggtitle("fill = 'lightblue'"),
          ggplot(auto, aes(acceleration)) + geom_histogram(fill = '#FFDB6D') + ggtitle("fill = #FFDB6D"),
          ggplot(auto, aes(acceleration)) + geom_histogram(linetype="dashed", fill = 'pink') +
            ggtitle("linetype='dashed', fill = 'pink'"),
          ggplot(auto, aes(acceleration)) + geom_histogram(color = 'red', fill = 'blue', linetype='dashed') +
            ggtitle("color = 'red', fill = 'blue', linetype='dashed'"),
          ncol = 2)
```


### change font size...
change font size:
```{r, fig.width=10,fig.height=11, message=FALSE}
h0 <- ggplot(auto,aes(acceleration)) + geom_histogram()
plot_grid(h0 + theme(text=element_text(size=20)) + ggtitle('text=element_text(size=20)'),
          h0 + ggtitle('text=element_text(size=NULL)'),
          h0 + theme(axis.text=element_text(size=20)) + ggtitle('axis.text=element_text(size=20)'),
          h0 + ggtitle('axis.text=element_text(size=NULL)'),
          h0 + theme(axis.title=element_text(size=20)) + ggtitle('axis.title=element_text(size=20)'),
          h0 + ggtitle('axis.title=element_text(size=NULL)'),
          h0 + theme(plot.title=element_text(size=20)) + ggtitle('plot.title=element_text(size=20)'),
          h0 + ggtitle('plot.title=element_text(size=NULL)'),
          ncol = 2)
```


### add title, rename labs...
now we will gonna add a title, change x and y names. We have two options:
1)
```{r, message=FALSE}
A <-  h0 + labs(x = 'x label', y = 'y label', title = 'A plot')
##
B <- h0 + xlab('x label') + ylab('y label') + ggtitle('B plot')
plot_grid(A,B)
```

other useful stuff with labs function:
2)
```{r}
h0 + labs(title = 'p title', subtitle = 'p subtitle', tag = 'p0')
```


### change axis limits.
change x axis limits:
```{r}
h0 +  scale_x_continuous(limits=c(min(auto$acceleration),max(auto$acceleration)), #future use: just change df$col
                         breaks=seq(min(auto$acceleration),max(auto$acceleration),1)) #you can change the by(100)
```


average line
```{r}
ggplot(auto, aes(acceleration)) + geom_histogram(fill = '#E69E80')  +
  geom_vline(aes(xintercept=mean(acceleration)), color="blue", linetype="dashed", size=1)
```


## Bivariate Histograms - with second categorical variable.
```{r}
plot_grid(ggplot(auto, aes(acceleration, color = origin)) + geom_histogram() + ggtitle('color'),
          ggplot(auto, aes(acceleration, fill = origin)) + geom_histogram() + ggtitle('fill') +
  scale_fill_brewer(palette="Dark2"))
```

### average line by group
```{r}
df <- auto %>% group_by(origin) %>% summarise(acceleration = mean(acceleration)) %>% as.data.frame()
```
```{r}
ggplot(data = auto, aes(acceleration, fill = origin)) + geom_histogram() + ggtitle('fill') +
  geom_vline(data = df, aes(xintercept=acceleration, color = origin), linetype="dashed", size=1) +
    scale_fill_grey() + scale_color_manual(values =  c('blue','yellow','green'))
```

### facet

#### facet grid
```{r,  fig.width=9,fig.height=5, message = FALSE}
plot_grid(
  ggplot(auto, aes(acceleration, fill = origin)) + geom_histogram() +
            ggtitle('facet_grid(origin ~ .)') + facet_grid(origin ~ .),
  
  ggplot(auto, aes(acceleration, color = origin)) + geom_histogram() + 
    ggtitle('facet_grid(. ~ origin)') + facet_grid(. ~ origin)
)
```

#### facet wrap
```{r,  fig.width=9,fig.height=5, message = FALSE}
plot_grid(
  ggplot(auto, aes(acceleration, fill = origin)) + geom_histogram() +
            ggtitle('facet_wrap(origin ~ .)') + facet_wrap(origin ~ .),
  
  ggplot(auto, aes(acceleration, color = origin)) + geom_histogram() + 
    ggtitle('facet_wrap( ~ origin)') + facet_wrap( ~ origin, ncol = 2)
)
```





```{r,  fig.width=6,fig.height=3, message=FALSE}
plot_grid(ggplot(auto,aes(acceleration)) + geom_boxplot() +  xlab("") +
            ggtitle("Histogram and Box-Plot") + theme(plot.title = element_text(hjust = 0.4)) +
            theme(aspect.ratio=1/10,  panel.border = element_blank()) + coord_fixed(),
          h0,
          ncol = 1)
```




Bivariate graph: showing the relationship between two variables.
```{r, message = FALSE}
GGally::ggpairs(select(auto,-c(origin,cylinders)))
```



