---
title: "EDA - Histograms"
author: "Khaled Sobeh"
date: "5/5/2022"
output: html_document
---

Data Visualization: Histograms / continuous data types 


Histogram is a kind of graphs that we use to represent a continuous variable.
1. By saying continuous, we mean a variable that can any value in an interval(such as: height, acceleration, length, time, temperature, age...), it's a quantitative variable that we can summarize.

Data structure in R: 
1. integer.
2. numeric.


I will try to use only ggplot2 functions, except if the alternative is much easier, with much less code to write. 
note: plot_grid() function will be used a lot. I used it just to combine plots, as we shall see.

So, let's start.

Loading the libraries, and getting the data.
```{r, message = FALSE, warning = FALSE}
library(ggplot2)
library(priceR)
library(dplyr)
library(forcats)
library(cowplot)
library(ISLR)
```

```{r}
auto <- Auto
auto$origin <- as.factor(auto$origin)
auto$cylinders <- as.factor(auto$cylinders)
auto <- select(auto, -name)
```
```{r}
str(auto)
```

The variables that I used from the dataset:
1) acceleration: continuous variable.

## a reminder.
If you forget one day how to make plot with ggplot2(Please don't...), then you have to remember that you have to pass these three:
1) ggplot(and_here_the_dataset_name)
2) aes(x = the_variable_name_that_you_want_to_visualize, y = sec_col...)
you can pass this aes thing(you should remember this) in ggplot() or geom_..() function.
3) geom_ the_plot_that_you_want: geom_bar(), geom_boxplot().....
so you should have:
ggplot(travel,aes(acceleration)) + geom_bar() |or| ggplot(travel) + geom_bar(aes(acceleration))


## The most basic histogram:
```{r}
ggplot(auto,aes(acceleration)) + geom_histogram()
```

we can also set the aesthetic in the geom function:
```{r}
ggplot(auto) + geom_histogram(aes(acceleration))
```

```{r, message=FALSE}
plot_grid(ggplot(auto,aes(acceleration)) + geom_histogram(),
          ggplot(auto) + geom_histogram(aes(acceleration)))
```


geom_histogram parameters:
```{r, fig.width=10,fig.height=11, message = FALSE}

plot_grid(ggplot(auto, aes(acceleration)) + geom_histogram(bins = 10) + ggtitle("bins = 10"),
          ggplot(auto, aes(acceleration)) + geom_histogram(bins = 50) + ggtitle("bins = 50"),
          ggplot(auto, aes(acceleration)) + geom_histogram(alpha = 0.3) + ggtitle("alpha = 0.3"),
          ggplot(auto, aes(acceleration)) + geom_histogram(alpha = 1) + ggtitle("alpha = 1"),
          ggplot(auto, aes(acceleration)) + geom_histogram(color = 'lightblue') + ggtitle("color = 'lightblue'"),
          ggplot(auto, aes(acceleration)) + geom_histogram(color = '#FFDB6D') + ggtitle("color = #FFDB6D"),
          ggplot(auto, aes(acceleration)) + geom_histogram(fill = 'lightblue') + ggtitle("fill = 'lightblue'"),
          ggplot(auto, aes(acceleration)) + geom_histogram(fill = '#FFDB6D') + ggtitle("fill = #FFDB6D"),
          #ggplot(auto, aes(acceleration)) + geom_histogram() + ggtitle(""),
          #ggplot(auto, aes(acceleration)) + geom_histogram() + ggtitle(""),
          ncol = 2)
```


### change font size...
change font size:
```{r, fig.width=10,fig.height=11, message=FALSE}
h0 <- ggplot(auto,aes(acceleration)) + geom_histogram()
plot_grid(h0 + theme(text=element_text(size=20)) + ggtitle('text=element_text(size=20)'),
          h0 + ggtitle('text=element_text(size=NULL)'),
          h0 + theme(axis.text=element_text(size=20)) + ggtitle('axis.text=element_text(size=20)'),
          h0 + ggtitle('axis.text=element_text(size=NULL)'),
          h0 + theme(axis.title=element_text(size=20)) + ggtitle('axis.title=element_text(size=20)'),
          h0 + ggtitle('axis.title=element_text(size=NULL)'),
          h0 + theme(plot.title=element_text(size=20)) + ggtitle('plot.title=element_text(size=20)'),
          h0 + ggtitle('plot.title=element_text(size=NULL)'),
          ncol = 2)
```


### add title, rename labs...
now we will gonna add a title, change x and y names. We have two options:
1)
```{r, message=FALSE}
A <-  h0 + labs(x = 'x label', y = 'y label', title = 'A plot')
##
B <- h0 + xlab('x label') + ylab('y label') + ggtitle('B plot')
plot_grid(A,B)
```

other useful stuff with labs function:
2)
```{r}
h0 + labs(title = 'p title', subtitle = 'p subtitle', tag = 'p0')
```


### change axis limits.
change x axis limits:
```{r}
h0 +  scale_x_continuous(limits=c(min(auto$acceleration),max(auto$acceleration)), #future use: just change df$col
                         breaks=seq(min(auto$acceleration),max(auto$acceleration),1)) #you can change the by(100)
```



Bivariate graph: showing the relationship between two variables.
```{r, message = FALSE}
GGally::ggpairs(select(auto,-c(origin,cylinders)))
```









