---
title: "DA - Box-Plots"
author: "Khaled Sobeh"
date: "13/5/2022"
output: html_document
---

Data Visualization: Box-Plots / continuous data types 


Box Plot is a kind of graphs that we use to represent a continuous variable.
1. By saying continuous, we mean a variable that can any value in an interval(such as: height, acceleration, length, time, temperature, age...), it's a quantitative variable that we can summarize.

Data structure in R: 
1. integer.
2. numeric.

Box-Plots
Box-Plots is visualization for a statistical measures of the data:
- Q1: the lower(first) quartile
- median
- Q3: the upper(third) quartile
- Maximum value: = Q3 + 1.5*IQR
- Minimum value: = Q1 - 1.5*IQR
- Outliers: data points Beyond Maximum/Minumum values

I will try to use only ggplot2 functions, except if the alternative is much easier, with much less code to write. 
note: plot_grid() function will be used a lot. I used it just to combine plots, as we shall see.

So, let's start.

Loading the libraries, and getting the data.
```{r, message = FALSE, warning = FALSE}
library(ggplot2)
library(priceR)
library(dplyr)
library(forcats)
library(cowplot)
library(ISLR)
```

```{r}
auto <- Auto
auto$origin <- as.factor(auto$origin)
auto$cylinders <- as.factor(auto$cylinders)
auto <- select(auto, -name)
```
```{r}
str(auto)
```


the most basic box-plot:
```{r, message=FALSE}
plot_grid(ggplot(auto,aes(acceleration)) + geom_boxplot(),
          ggplot(auto) + geom_boxplot(aes(acceleration)))
```

```{r}
summary(auto$acceleration)
```


rotate the box-plot:
```{r, message=FALSE}
plot_grid(ggplot(auto,aes(y = acceleration)) + geom_boxplot(),
          ggplot(auto,aes(x = acceleration)) + geom_boxplot())
```

geom_boxplot parameters:
```{r, fig.width=10,fig.height=11, message = FALSE}
plot_grid(
          ggplot(auto, aes(acceleration)) + geom_boxplot(alpha = 0.3) + ggtitle("alpha = 0.3"),
          ggplot(auto, aes(acceleration)) + geom_boxplot(alpha = 1) + ggtitle("alpha = 1"),
          ggplot(auto, aes(acceleration)) + geom_boxplot(color = 'lightblue') + ggtitle("color = 'lightblue'"),
          ggplot(auto, aes(acceleration)) + geom_boxplot(color = '#FFDB6D') + ggtitle("color = #FFDB6D"),
          ggplot(auto, aes(acceleration)) + geom_boxplot(fill = 'lightblue') + ggtitle("fill = 'lightblue'"),
          ggplot(auto, aes(acceleration)) + geom_boxplot(fill = '#FFDB6D') + ggtitle("fill = #FFDB6D"),
          #ggplot(auto, aes(acceleration)) + geom_boxplot() + ggtitle(""),
          #ggplot(auto, aes(acceleration)) + geom_boxplot() + ggtitle(""),
          ncol = 2)
```

add the average:
```{r}
ggplot(auto,aes(x = acceleration, y = factor(0))) + geom_boxplot() +
  #stat_summary(fun.x="mean")
  stat_summary(fun.x=mean, geom="point", shape=20, size=14, color="red", fill="red")
```




```{r}
ggplot(auto,aes(x = acceleration)) + geom_boxplot() + geom_dotplot(binwidth = 1/10)
```


