---
title: "DA - Density Plots"
author: "Khaled Sobeh"
date: "5/5/2022"
output: html_document
---

Data Visualization: Density Plots / continuous data types 


Density Plots is a kind of graphs that we use to represent a continuous variable.
1. By saying continuous, we mean a variable that can any value in an interval(such as: height, acceleration, length, time, temperature, age...), it's a quantitative variable that we can summarize.

Data structure in R: 
1. integer.
2. numeric.

Density Plots
- with Density Plots we can check if the "pattern" of a variable ,if the variable is: "symmetric", "skewed left" or "right", "unimodal"(normally distributed), "bimodal" or "multimodal". 
- a nice chose when we have too many data.

How it's work?
1. put each data point, and add a Gaussian density kernel for each one.
2. sum the kernels together.

I will try to use only ggplot2 functions, except if the alternative is much easier, with much less code to write. 
note: plot_grid() function will be used a lot. I used it just to combine plots, as we shall see.

So, let's start.

Loading the libraries, and getting the data.
```{r, message = FALSE, warning = FALSE}
library(ggplot2)
library(priceR)
library(dplyr)
library(forcats)
library(cowplot)
library(ISLR)
```

```{r}
auto <- Auto
auto$origin <- as.factor(auto$origin)
auto$cylinders <- as.factor(auto$cylinders)
auto <- select(auto, -name)
```
```{r}
str(auto)
```


```{r, message=FALSE}
plot_grid(ggplot(auto,aes(acceleration)) + geom_density(),
          ggplot(auto) + geom_density(aes(acceleration)))
```

Important settings in density plots:
```{r}
plot_grid(ggplot(auto,aes(acceleration)) + geom_density(bw = 5) + ggtitle('bw = 5'),
          ggplot(auto,aes(acceleration)) + geom_density(bw = 15) + ggtitle('bw = 15'))
```

```{r}
ggplot(auto,aes(acceleration)) + geom_density() + geom_rug()
```

geom_density parameters: alpha, color, fill, group, linetype, size, weight
```{r, fig.width=10,fig.height=11, message = FALSE}

plot_grid(ggplot(auto, aes(acceleration)) + geom_density(alpha = 0.3) + ggtitle("alpha = 0.3"),
          
          ggplot(auto, aes(acceleration)) + geom_density(alpha = 1, size = 1) + ggtitle("alpha = 1, size = 1"),
          
          ggplot(auto, aes(acceleration)) + geom_density(color = 'lightblue', size = 1) + ggtitle("color = 'lightblue', size = 1"),
          
          ggplot(auto, aes(acceleration)) + geom_density(color = '#FFDB6D', size = 1) + ggtitle("color = #FFDB6D, size = 1"),
          
          ggplot(auto, aes(acceleration)) + geom_density(fill = 'lightblue') + ggtitle("fill = 'lightblue'"),
          
          ggplot(auto, aes(acceleration)) + geom_density(fill = 'green') + ggtitle("fill = green"),
          
          ggplot(auto, aes(acceleration)) + geom_density(linetype = "dashed",color="darkblue",size = 1) +
            ggtitle("linetype = 'dashed', color='darkblue', size = 1"),
          #ggplot(auto, aes(acceleration)) + geom_density() + ggtitle(""),
          ncol = 2)
```


add mean line:
```{r}
ggplot(auto, aes(acceleration)) + geom_density(alpha = 1, size = 1, fill = '#E69E80')  +
  geom_vline(aes(xintercept=mean(acceleration)), color="blue", linetype="dashed", size=1)
```








```{r}
ggplot(auto, aes(x=weight)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill = "cyan3") 
```





References:
https://stat-lu.github.io/dataviz/
http://www.sthda.com/english/wiki/ggplot2-density-plot-quick-start-guide-r-software-and-data-visualization