---
title: "DA - Box-Plots"
author: "Khaled Sobeh"
date: "13/5/2022"
output: html_document
---

Data Visualization: Box-Plots / continuous data types 


Box Plot is a kind of graphs that we use to represent a continuous variable.
1. By saying continuous, we mean a variable that can any value in an interval(such as: height, acceleration, length, time, temperature, age...), it's a quantitative variable that we can summarize.

Data structure in R: 
1. integer.
2. numeric.

Box-Plots
Box-Plots is a visualization for statistical measures of the data:
- Q1: the lower(first) quartile
- median(Q2)
- Q3: the upper(third) quartile
- Maximum value: = Q3 + 1.5*IQR
- Minimum value: = Q1 - 1.5*IQR
- Outliers: data points Beyond Maximum/Minumum values
note: IQR = Q3 - Q1
```{r, echo=FALSE}
auto <- ISLR::Auto
sum_acc <- summary(auto$acceleration)
```
```{r, echo=FALSE}
ggplot(auto) + 
  geom_boxplot(aes(y = acceleration, x = factor(0))) + 
  annotate(geom="text", x=factor(0), y=sum_acc[3] + 0.5, 
           label="median", color="#E69F00") +
  annotate(geom="text", x=factor(0), y=sum_acc[2] + 0.5,
           label="1st Quartile                     ", color="#E69F00") +
  annotate(geom="text", x=factor(0), y=sum_acc[5]+ 0.5,
           label="3rd Quartile                     ", color="#E69F00") +
  annotate(geom="text", x=factor(0) , y=(sum_acc[2] - 1.5*(sum_acc[5]-sum_acc[2])) + 0.5,
           label="Q1 - 1.5*IQR                          ", color="#E69F00") +
  annotate(geom="text", x=factor(0) , y=(sum_acc[2] - 1.5*(sum_acc[5]-sum_acc[2])) + 0.5,
           label="_____________                         ", color="#E69F00") +
  annotate(geom="text", x=factor(0), y=(sum_acc[5] + 1.5*(sum_acc[5]-sum_acc[2])) + 0.2,
           label="Q3 + 1.5*IQR                          ", color="#E69F00") +
  annotate(geom="text", x=factor(0), y=(sum_acc[5] + 1.5*(sum_acc[5]-sum_acc[2])) + 0.2,
           label="_____________                         ", color="#E69F00") +
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
  theme(axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) + 
  ggtitle("Box-Plot:")
```


I will try to use only ggplot2 functions, except if the alternative is much easier, with much less code to write. 
note: plot_grid() function will be used a lot. I used it just to combine plots, as we shall see.

So, let's start.

Loading the libraries, and getting the data.
```{r, message = FALSE, warning = FALSE}
library(ggplot2)
library(priceR)
library(dplyr)
library(forcats)
library(cowplot)
library(ISLR)
```

```{r}
auto <- Auto
auto$origin <- as.factor(auto$origin)
auto$cylinders <- as.factor(auto$cylinders)
auto <- select(auto, -name)
```
```{r}
str(auto)
```


the most basic box-plot:
```{r, message=FALSE}
plot_grid(ggplot(auto,aes(acceleration)) + geom_boxplot(),
          ggplot(auto) + geom_boxplot(aes(acceleration)))
```

```{r}
summary(auto$acceleration)
```


rotate the box-plot:
```{r, message=FALSE}
plot_grid(ggplot(auto,aes(y = acceleration)) + geom_boxplot(),
          ggplot(auto,aes(x = acceleration)) + geom_boxplot(),
          qplot(y = acceleration,data =  auto, geom="boxplot"),
          qplot(x = acceleration,data =  auto, geom="boxplot"),
          ncol = 2)
```

geom_boxplot parameters:
```{r, fig.width=10,fig.height=11, message = FALSE}
plot_grid(
          ggplot(auto, aes(acceleration)) + geom_boxplot(alpha = 0.3) + ggtitle("alpha = 0.3"),
          ggplot(auto, aes(acceleration)) + geom_boxplot(alpha = 1) + ggtitle("alpha = 1"),
          ggplot(auto, aes(acceleration)) + geom_boxplot(color = 'lightblue') + ggtitle("color = 'lightblue'"),
          ggplot(auto, aes(acceleration)) + geom_boxplot(color = '#FFDB6D') + ggtitle("color = #FFDB6D"),
          ggplot(auto, aes(acceleration)) + geom_boxplot(fill = 'lightblue') + ggtitle("fill = 'lightblue'"),
          ggplot(auto, aes(acceleration)) + geom_boxplot(fill = '#FFDB6D') + ggtitle("fill = #FFDB6D"),
          #ggplot(auto, aes(acceleration)) + geom_boxplot() + ggtitle(""),
          #ggplot(auto, aes(acceleration)) + geom_boxplot() + ggtitle(""),
          ncol = 2)
```

add the average:
and shape, size
```{r}
plot_grid(ggplot(auto,aes(x = acceleration, y = factor(0))) + geom_boxplot() + labs(tag = 'A') +
  stat_summary(fun.x=mean, geom="point", shape=23, size=7, fill="lightblue"),
  
  ggplot(auto,aes(y = acceleration, x = 0)) + geom_boxplot() + labs(tag = 'B') +
  stat_summary(fun.y=mean, geom="point", shape=20, size=5, color="red", fill="red"),
  
  ggplot(auto,aes(y = acceleration, x = 0)) + geom_boxplot() + labs(tag = 'C') +
  stat_summary(fun.y=mean, geom="point", shape=22, size=5, color="red", fill="black"),
  
  ggplot(auto,aes(y = acceleration, x = 0)) + geom_boxplot() + labs(tag = 'D') +
  stat_summary(fun.y=mean, geom="point", shape=2, size=5, color="lightgreen", fill="red"))
```





```{r}
ggplot(auto,aes(x = acceleration)) + geom_boxplot() + geom_dotplot(binwidth = 1/10)
```

```{r}
ggplot(auto,aes(x = acceleration)) + geom_boxplot() + geom_dotplot(binwidth = 1/10) + coord_flip()
```


```{r}
ggplot(auto,aes(x = acceleration, y = factor(0))) + geom_boxplot() +
  geom_dotplot(binwidth = 1/5)  +
  stat_summary(fun.x=mean, geom="point", shape=23, size=5, fill="lightblue")
```



, stackdir = "center"

```{r, message=FALSE, warning=FALSE}
ggplot(auto) + 
  geom_boxplot(aes(x = acceleration, y = factor(0))) +
  geom_dotplot(aes(x = acceleration), binwidth = 1/5, alpha = 0.2, stackdir = "down")  + 
  geom_density(aes(x = acceleration), size = 1) + 
  stat_summary(aes(x = acceleration, y = factor(0)),fun.x=mean, geom="point", shape=23, size=3, fill="lightblue") +
  stat_summary(aes(x = acceleration, y = factor(0), label = paste("average=", round(..x.., digits = 1))), fun.x = mean, geom = "text", col = "blue", vjust = 2.5) +
  theme(axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank())
```


```{r}
plot_continuous <- function(df, xcol){
  ggplot(df) + 
    geom_boxplot(aes(x = xcol, y = factor(0))) +
  geom_dotplot(aes(x = xcol), binwidth = 1/5, alpha = 0.2, stackdir = "down")  + 
  geom_density(aes(x = xcol), size = 1) + 
  stat_summary(aes(x = xcol, y = factor(0)),fun.x=mean, geom="point", shape=23, size=3, fill="lightblue") +
  stat_summary(aes(x = xcol, y = factor(0), label = paste("average=", round(..x.., digits = 1))), fun.x = mean, geom = "text", col = "blue", vjust = 2.5) +
  theme(axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) +
    coord_cartesian(ylim = c(-1,1))
}
```
```{r}
plot_continuous(auto,auto$mpg)
```


```{r}
ggplot(auto) + 
  geom_histogram(aes(x = mpg)) +
  #geom_dotplot(aes(x = mpg), alpha = 0.2)  + 
  geom_density(aes(x = mpg), size = 1)
```



different types of boxplot:
```{r}
ggplot(auto, aes(y = mpg)) + geom_boxplot(notch = TRUE) + ggtitle('notch = TRUE')
```


for a few data you can use:
```{r}
ggplot(auto, aes(x = mpg)) + geom_dotplot(binwidth = 1/1.5) + ggtitle('dotplot') + guides(y = "none")
```

for a lot data you can use:
```{r}
ggplot(auto, aes(y = mpg, x = 1)) + geom_violin() + ggtitle('violin')
```








References:
http://www.sthda.com/english/wiki/be-awesome-in-ggplot2-a-practical-guide-to-be-highly-effective-r-software-and-data-visualization





