---
title: "DA - Box-Plots"
author: "Khaled Sobeh"
date: "13/5/2022"
output: html_document
---

Data Visualization: Box-Plots / continuous data types 


Box Plot is a kind of graphs that we use to represent a continuous variable.
1. By saying continuous, we mean a variable that can any value in an interval(such as: height, acceleration, length, time, temperature, age...), it's a quantitative variable that we can summarize.

Data structure in R: 
1. integer.
2. numeric.

Box-Plots
Box-Plots is a visualization for statistical measures of the data:
- Q1: the lower(first) quartile
- median(Q2)
- Q3: the upper(third) quartile
- Maximum value: = Q3 + 1.5*IQR
- Minimum value: = Q1 - 1.5*IQR
- Outliers: data points Beyond Maximum/Minumum values
note: IQR = Q3 - Q1
```{r, echo=FALSE}
auto <- ISLR::Auto
sum_acc <- summary(auto$acceleration)
```
```{r, echo=FALSE}
ggplot2::ggplot(auto) + 
  geom_boxplot(aes(y = acceleration, x = 1), outlier.alpha = 1, outlier.color = 'red') + 
  annotate(geom="text", x=1, y=sum_acc[3] + 0.5, # text for median
           label="median", color="#E69F00", family = "serif", fontface = "italic", size = 5) +
  
  annotate("segment", # outliers down
           y = (sum_acc[2] - 1.5*(sum_acc[5]-sum_acc[2])),
           yend = min(auto$acceleration),
           x = 1.02,xend = 1.02, color="brown",size =2,
           arrow = arrow( length = unit(.2,"cm"))) +
  annotate(geom="text", x=1.033, y=sum_acc[6]-1.5, 
           label="Outliers", color="red", angle = 90) +
  
  
  annotate("segment", # outliers up
           y = (sum_acc[5] + 1.5*(sum_acc[5]-sum_acc[2])),
           yend = max(auto$acceleration),
           x = 1.02,xend = 1.02, color="brown",size =2,
           arrow = arrow( length = unit(.2,"cm"))) +
  annotate(geom="text", x=1.06, y=sum_acc[1]+0.5, 
           label="Outliers", color="red") +
  
  annotate(geom="text", x=1.4, y=sum_acc[3], family = "serif", fontface = "italic", size = 6, #IQR text and arrow
           label="IQR", color="blue", angle = 90) +
  annotate("segment", y = sum_acc[2], yend = sum_acc[5],x = 1.39,xend = 1.39, color="brown", size = 1, 
           arrow = arrow(ends = "both", angle = 90, length = unit(.2,"cm"))) +
  
  
  annotate(geom="text", x=1, y=sum_acc[2] + 0.5, family = "serif", fontface = "italic", size = 5,
           label="1st Quartile                     ", color="#E69F00") +
  annotate(geom="text", x=1, y=sum_acc[5]+ 0.5, family = "serif", fontface = "italic", size = 5,
           label="3rd Quartile                     ", color="#E69F00") +
  annotate(geom="text", x=1 , y=(sum_acc[2] - 1.5*(sum_acc[5]-sum_acc[2])) + 0.5,
           family = "serif", fontface = "italic", size = 5,
           label="Q1 - 1.5*IQR                          ", color="#E69F00") +
  annotate(geom="text", x=1 , y=(sum_acc[2] - 1.5*(sum_acc[5]-sum_acc[2])) + 0.5,
           family = "serif", fontface = "italic", size = 5,
           label="____________                          ", color="#E69F00") +
  annotate(geom="text", x=1, y=(sum_acc[5] + 1.5*(sum_acc[5]-sum_acc[2])) + 0.5,
           family = "serif", fontface = "italic", size = 5,
           label="Q3 + 1.5*IQR                          ", color="#E69F00") +
  annotate(geom="text", x=1, y=(sum_acc[5] + 1.5*(sum_acc[5]-sum_acc[2])) + 0.5,
           family = "serif", fontface = "italic", size = 5,
           label="____________                          ", color="#E69F00") +
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
  theme(axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) + 
  ggtitle("Box-Plot")

```


I will try to use only ggplot2 functions, except if the alternative is much easier, with much less code to write. 
note: plot_grid() function will be used a lot. I used it just to combine plots, as we shall see.

So, let's start.

Loading the libraries, and getting the data.
```{r, message = FALSE, warning = FALSE}
library(ggplot2)
library(priceR)
library(dplyr)
library(forcats)
library(cowplot)
library(ISLR)
```

```{r}
auto <- Auto
auto$origin <- as.factor(auto$origin)
auto$cylinders <- as.factor(auto$cylinders)
auto <- select(auto, -name)
```
```{r}
str(auto)
```


the most basic box-plot:
```{r, message=FALSE}
plot_grid(ggplot(auto,aes(acceleration)) + geom_boxplot(),
          ggplot(auto) + geom_boxplot(aes(acceleration)))
```

```{r}
summary(auto$acceleration)
```


rotate the box-plot:
```{r, message=FALSE}
plot_grid(ggplot(auto,aes(y = acceleration)) + geom_boxplot(),
          ggplot(auto,aes(x = acceleration)) + geom_boxplot(),
          qplot(y = acceleration,data =  auto, geom="boxplot"),
          qplot(x = acceleration,data =  auto, geom="boxplot"),
          ncol = 2)
```

geom_boxplot parameters:
```{r, fig.width=10,fig.height=11, message = FALSE}
plot_grid(
          ggplot(auto, aes(acceleration)) + geom_boxplot(alpha = 0.3) + ggtitle("alpha = 0.3"),
          ggplot(auto, aes(acceleration)) + geom_boxplot(alpha = 1) + ggtitle("alpha = 1"),
          ggplot(auto, aes(acceleration)) + geom_boxplot(color = 'lightblue') + ggtitle("color = 'lightblue'"),
          ggplot(auto, aes(acceleration)) + geom_boxplot(color = '#FFDB6D') + ggtitle("color = #FFDB6D"),
          ggplot(auto, aes(acceleration)) + geom_boxplot(fill = 'lightblue') + ggtitle("fill = 'lightblue'"),
          ggplot(auto, aes(acceleration)) + geom_boxplot(fill = '#FFDB6D') + ggtitle("fill = #FFDB6D"),
          #ggplot(auto, aes(acceleration)) + geom_boxplot() + ggtitle(""),
          #ggplot(auto, aes(acceleration)) + geom_boxplot() + ggtitle(""),
          ncol = 2)
```

add the average:
and shape, size
```{r}
plot_grid(ggplot(auto,aes(x = acceleration, y = factor(0))) + geom_boxplot() + labs(tag = 'A') +
  stat_summary(fun.x=mean, geom="point", shape=23, size=7, fill="lightblue"),
  
  ggplot(auto,aes(y = acceleration, x = 0)) + geom_boxplot() + labs(tag = 'B') +
  stat_summary(fun.y=mean, geom="point", shape=20, size=5, color="red", fill="red"),
  
  ggplot(auto,aes(y = acceleration, x = 0)) + geom_boxplot() + labs(tag = 'C') +
  stat_summary(fun.y=mean, geom="point", shape=22, size=5, color="red", fill="black"),
  
  ggplot(auto,aes(y = acceleration, x = 0)) + geom_boxplot() + labs(tag = 'D') +
  stat_summary(fun.y=mean, geom="point", shape=2, size=5, color="lightgreen", fill="red"))
```





```{r}
plot_grid(
  ggplot(auto,aes(x = acceleration)) + geom_boxplot() + geom_dotplot(binwidth = 1/10),
ggplot(auto,aes(x = acceleration)) + geom_boxplot() + geom_dotplot(binwidth = 1/10) + coord_flip()
)
```


```{r}
ggplot(auto,aes(x = acceleration, y = factor(0))) + geom_boxplot() +
  geom_dotplot(binwidth = 1/5)  +
  stat_summary(fun.x=mean, geom="point", shape=23, size=5, fill="lightblue")
```



, stackdir = "center"

```{r, message=FALSE, warning=FALSE}
ggplot(auto) + 
  geom_boxplot(aes(x = acceleration, y = factor(0))) +
  geom_dotplot(aes(x = acceleration), binwidth = 1/5, alpha = 0.2, stackdir = "down")  + 
  geom_density(aes(x = acceleration), size = 1) + 
  stat_summary(aes(x = acceleration, y = factor(0)),fun.x=mean, geom="point", shape=23, size=3, fill="lightblue") +
  stat_summary(aes(x = acceleration, y = factor(0), label = paste("average=", round(..x.., digits = 1))), fun.x = mean, geom = "text", col = "blue", vjust = 2.5) +
  theme(axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank())
```


## notched boxplot
The notch displays a confidence interval around the median which is normally based on the median +/- 1.58*IQR/sqrt(n). 
Notches are used to compare groups; if the notches of two boxes do not overlap, this is a strong evidence that the medians differ.
```{r}
ggplot(auto, aes(y = mpg)) + geom_boxplot(notch = TRUE) + ggtitle('notch = TRUE')
```

jitter plot
```{r}
ggplot(auto,aes(x = acceleration, y = 1)) + geom_jitter(shape=16, position=position_jitter(0.2)) +
  theme(
    text = element_text(colour = "red"),
    axis.title = element_text(size = rel(1.25))
  )
```

```{r}
plot_grid(ggplot(auto,aes(x = acceleration, y = 1)) + 
            geom_jitter(shape=16, position=position_jitter(0.2)),
          
          ggplot(auto,aes(x = acceleration, y = 1)) + geom_boxplot() +
            geom_jitter(shape=16, position=position_jitter(width = 0.1)))
```



```{r}
ggplot(auto, aes(x = mpg)) + geom_dotplot(binaxis='x', stackdir='center', dotsize=0.5) +
  ylab("Count\n(the mpg)") +
  theme(axis.title.y = element_text( angle = 90,face = "italic", colour = "darkred",size = 14),
  axis.title.x = element_text(face = "italic", colour = "darkred", size = 14)) 
  
```


dot-plot:

for a few data you can use:
```{r}
#extrafont::loadfonts(device = 'win')
ggplot(auto, aes(x = mpg)) + geom_dotplot(binwidth = 1/1.5) +  guides(y = "none") +
  ggtitle("dotplot \nfor small data") +
  theme(plot.title = element_text(
    size = rel(1.5), lineheight = .9,
    family = "Times", face = "bold.italic", colour = "red"
  ))
```

# Violin Plot
- It shows the probability density of the data, using smoothed kernel density estimatior.
- for continuous data.

for a lot data you can use:
```{r, fig.width=2.5,fig.height=5}
ggplot(auto, aes(y = mpg, x = 1)) + geom_violin() + ggtitle('violin')
```

qplot
```{r}
ggplot(auto, aes(sample = acceleration)) +
  geom_qq() +
  geom_qq_line()
```

```{r}
ggplot(auto, aes(acceleration)) + stat_ecdf()
#m <- lm(y2 ~ x3 + x4)
#par(mfrow = (c(2, 2)))      # this gives us a 2x2 plot
#plot(m)
```


```{r}
ggplot(auto, aes(y = acceleration, x = 1)) + geom_violin() +
    geom_boxplot(width=0.1, color="grey", alpha=0.2, width = 0.1,
                 outlier.alpha = 1, outlier.color = 'red') 
```

```{r}
ggplot(auto, aes(y = acceleration, x = 1)) + gghalves::geom_half_violin() +
  gghalves::geom_half_boxplot(alpha=0.2, width = 0.1, outlier.alpha = 1, outlier.color = 'red') 
```


```{r}
ggplot(auto, aes(y = acceleration, x = 1)) + gghalves::geom_half_violin() +
  gghalves::geom_half_boxplot(alpha=0.2, width = 0.1, outlier.alpha = 1, outlier.color = 'red')  +
  gghalves::geom_half_dotplot(binwidth = 1/3)
```


```{r ,warning=FALSE, message=FALSE}
plot_grid(
  ggplot(auto,aes(acceleration)) + geom_histogram(),
  
  ggplot(auto, aes(y = acceleration, x = 1)) + gghalves::geom_half_violin() +
  #gghalves::geom_half_dotplot(binwidth = 1/8) +
  gghalves::geom_half_boxplot(alpha=0.2, width = 0.3, outlier.alpha = 1, outlier.color = 'red') +
    stat_summary(fun.y=mean, geom="point", shape=20, size=5, color="lightblue") +
    coord_flip() + labs(x = "", y = ""),
  
  ncol = 1
)
```


### Bivariate graphs: adding a second categorical variable.

```{r}
ggplot(auto, aes(y = acceleration, x = as.factor(origin))) + gghalves::geom_half_violin() +
  gghalves::geom_half_boxplot(alpha=0.2, width = 0.1, outlier.alpha = 1, outlier.color = 'red')  +
  gghalves::geom_half_dotplot(binwidth = 1/3) +
  theme(axis.text.x = element_text(family = "Times", face = "italic",
                               colour = "darkred", size = rel(0.9))) +
  scale_y_continuous(name = "Origin\n(as factor)")
```

```{r}
df <- auto %>% group_by(origin) %>% summarise(acceleration_mean = mean(acceleration),
                                              acceleration_max = max(acceleration),
                                              acceleration_min = min(acceleration)) %>%
  as.data.frame()
ggplot(auto,aes(x = acceleration, y = 1 ,fill =  origin)) + 
  geom_dotplot(binwidth = 1/3) + guides(y = "none") +
  scale_x_continuous(breaks = seq(min(auto$acceleration), max(auto$acceleration), by = 1)) +
  geom_vline(data = df, aes(xintercept=acceleration_min, color = origin), linetype="dashed", size=1) +
  geom_vline(data = df, aes(xintercept=acceleration_max, color = origin), linetype="dashed", size=1) +
  facet_grid(origin ~ .)

```




References:
http://www.sthda.com/english/wiki/be-awesome-in-ggplot2-a-practical-guide-to-be-highly-effective-r-software-and-data-visualization
http://www.sthda.com/english/wiki/ggplot2-box-plot-quick-start-guide-r-software-and-data-visualization

